---
title: "PCA on the Beer Data"
author: "Vincent Guillemot, Hervé Abdi, Ju-Chi Yu & Soudeh Ardestani Khoubrouy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PCA on the Beer Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type="text/css">
body {
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, 
  dev = "svg",
  fig.ext = "svg", 
  fig.width = 7, 
  fig.height = 7
)
library(dplyr)
library(ExPosition)
library(knitr)
library(kableExtra)
library(ggplot2)
library(InPosition)
set.seed(564)
```

## Load the data and the packages

With the following command

```{r beers, warning=FALSE, message=FALSE}
library(R4SPISE2022)
data("sixBeers12Descriptors10Judges", package = "data4PCCAR")
beers <- sixBeers12Descriptors10Judges$ratingsIntensity
beerColors <- sixBeers12Descriptors10Judges$color4Products
descColors <- sixBeers12Descriptors10Judges$color4Descriptors
```

 * `beers`: the ratings;
 * `beerColors`: colors for the beers;
 * `descColors`: colors for the descriptors.

## Do the Principal Component Analysis (PCA)

```{r epPCA}
res.pca <- epPCA(beers, 
                 center = TRUE, # Center the data
                 scale = FALSE, # **Do not** scale the variables
                 graphs = FALSE) 
```

The default of this function will center (to have means equal 0) and scale (to have the _standard deviations_ equal 1) all variables.

## Plot the results

```{r plotPCA, fig.show = 'hide'}
res.plot.pca <- OTAplot(
    resPCA = res.pca,
    data = beers
    )
```

 * `results.stats`: useful statistics after performing PCA, among which one can find the factor scores, loadings, eigenvalues, and contributions;
 * `results.graphs`: all the PCA graphs (heatmaps, screeplot, factor map, correlation circles, loading maps etc.) 
 * `description.graphs`: the titles of these graphs, used as titles in the PPTX file.
 
### Correlations of the variables

We can check the data by plotting the correlation matrix between the variables. This correlation matrix is where the components are extracted.

```{r heatmat cor}
res.plot.pca$results.graphs$correlation
```

### Scree plot

The scree plot shows the eigenvalues of each component. These eigenvalues give the _variance_ of each component (or called dimension in the figure). In other words, the singular values, which are the square root of the eigenvalues, give the _standard deviations_ of these components. The sum of the eigenvalues is equal to the total inertia of the data.

```{r scree}
res.plot.pca$results.graphs$scree
```

### Factor scores (rows)

Here, the observation map shows how the observations (i.e., rows) are represented in the component space. If `center = TRUE` when running PCA, the origin should be at the center of all observations. (If it's not, something is off.)

```{r factor scores}
res.plot.pca$results.graphs$factorScoresI12
```


### Circle of correlation

```{r correlation circle}
res.plot.pca$results.graphs$cosineCircleArrowJ12
```

## Inference plots and results

```{r plotPCAinference, fig.show = 'hide', warning = FALSE, message = FALSE}
res.plot.pca.inference <- OTAplotInference(
    resPCA = res.pca,
    data = beers
    ) %>% suppressMessages()
```


```{r scree inference, warning = FALSE, message = FALSE}
res.plot.pca.inference$results.graphs$scree
```

```{r bootstrap ratios 1}
res.plot.pca.inference$results.graphs$BR1
```

```{r bootstrap ratios 2}
res.plot.pca.inference$results.graphs$BR2
```

