---
title: "PLSC on Wines"
author: "Vincent Guillemot, Hervé Abdi, Ju-Chi Yu & Soudeh Ardestani Khoubrouy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PLSC on Wines}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
<style type="text/css">
body {
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, 
  dev = "svg",
  fig.ext = "svg", 
  fig.width = 7, 
  fig.height = 7
)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)
library(TExPosition)
```

## Load the course library

```{r library, warning = FALSE, message = FALSE}
library(R4SPISE2022)
```

## Load the data

With the following command

```{r winesOf3Colors}
data("winesOf3Colors", package = "data4PCCAR")
```

```{r hidden colors, echo = FALSE}
wineColors <- list()
wineColors$oc <- as.matrix(as.character(recode(winesOf3Colors$winesDescriptors$color, red = 'indianred4', 
                white = 'gold', rose = 'lightpink2')))
wineColors$gc <- as.matrix(c(red = 'indianred4', 
                white = 'gold', rose = 'lightpink2'))

varColors <- list()
varColors$oc[[1]] <- as.matrix(rep("darkorange1", 4))
varColors$oc[[2]] <- as.matrix(rep("olivedrab3", 8))
    
```

## Two tables with descriptors and other supplementary information

```{r four table, echo = TRUE}
descr <- winesOf3Colors$winesDescriptors %>% 
    select(origin, color, varietal)
suppl <- winesOf3Colors$winesDescriptors %>% 
    select(Price)
chemi <- winesOf3Colors$winesDescriptors %>% 
    select(Acidity, Alcohol, Sugar, Tannin)
senso <- winesOf3Colors$winesDescriptors %>% 
    select(fruity, floral, vegetal, 
           spicy, woody, sweet, astringent, 
           hedonic)
```

### Two data tables

 * First table: chemical data (`chemi`) that includes `acidity`, `alcohol`, `sugar`, and `tannin`.
 * Second table: sensory data (`senso`) that includes `fruity`, `floral`, `vegetal`, `spicy`, `woody`, `sweet`, `astringent`, and `henodic`.

```{r chemi, eval = FALSE, echo = FALSE, results='asis'}
### First table: chemical data

# fontsize <- 9.

# chemi %>%
#     kable("html", caption = "Table 1: Chemical data") %>%
#     kable_styling("striped", position = "float_left", full_width = F)  %>%
#     row_spec(1:nrow(chemi), 
#     font_size = fontsize) %>%
#     row_spec(0, angle = 0, align = "right", font_size = fontsize)
```


```{r senso, eval = FALSE, echo = FALSE}
### Second table: sensory data
# senso %>%
#     kable("html", caption = "Table 2: Sensory data") %>%
#     kable_styling("striped", position = "left", full_width = F)  %>%
#     row_spec(1:nrow(senso), 
#     font_size = fontsize) %>%
#     row_spec(0, angle = 0, align = "right", font_size = fontsize)
```

### Descriptors and supplementary variable

 * Descriptors: `origin`, `color` and `varietal`;
 * Suplementary variables: `Price`;
 
## Run PLS-C on the two tables

```{r run}
res.pls <- tepPLS(chemi, senso, DESIGN = descr$color, graphs = FALSE)
```

The default of this function will center (to have means equal 0) and scale (to have the _sums of squares_ equal 1) all variables in both data tables. The argument `DESIGN` indicates the groups of the observations and change how the observations are colored in the figures (when `graphs = TRUE`), but it does not change the results of PLS-C.

## Generate the figures

```{r TTA plots, fig.show='hide', message = FALSE, warning = FALSE}
## You might need to load these packages if the error is saying that it couldn't find some functions
# library(PTCA4CATA) 
# library(data4PCCAR)
res.pls.plot <- TTAplot(
    res = res.pls, # Output of tepPLS
    color.obs = wineColors, # <optional> colors of wines
    color.tab = varColors, # <optional> colors of the two tables
    tab1.name = "Chemical data", # <optional> Name of Table 1 (for printing)
    tab2.name = "Sensory data", # <optional> Name of Table 2 (for printing)
    DESIGN = descr$color, # design for the wines
    tab1 = chemi,  # First data table
    tab2 = senso)  # Second data table
```

In this `TTAplot` function, if `DESIGN` is specified. The latent variables will be colored according to the groups of the observations with the group means and their 95% bootstrap confidence intervals.

### Correlation between the two tables

We can check the data by plotting first the correlation matrix between the two data sets. This correlation matrix is where the dimensions are extracted.

```{r hetamap.rxy}
res.pls.plot$results.graphs$heatmap.rxy
```

### Scree plot

The scree plot shows the eigenvalues of each dimension. These eigenvalues give the squared _covariance_ of each pair of latent variables. In other words, the singular values, which are the square root of the eigenvalues, give the _covariance_ of these pairs of latent variables. The sum of the eigenvalues is equal to the sum of the squared covariance between all variables in both tables.

```{r scree, fig.height = 4}
res.pls.plot$results.graphs$scree.eig
```

### Latent variables

Here, we plot the first latent variable of both tables against each other with the observations colored according to their groups. This plot shows how the observations are distributed on the dimension and how the chosen pair of latent variables are related to each other. When plotting the first pair of latent variables, we expect the _observations_ to distribute along the bottom-left-to-top-right diagonal line (which illustrates a perfect correlation), because PLS-C maximizes the _covariance_ of the latent variables.

To examine the stability of these groups, we plot the group means with their 95% bootstrap confidence intervals (or ellipsoids). If the ellipses do not overlap, the groups are reliably different from each other. However, it's worth noted that the distribution of the _observations_ does not imply how the _groups_ (represented by the group means) are distributed or whether the groups are reliably different from each other. 

_Note_: The grouping information are independent from PLS-C and are only use to help provide a summary description of the observations.

```{r LV map}
res.pls.plot$results.graphs$lv.plot
```

The results from Dimension 1 show that the association between the chemical and the sensory data reliably separates the red wines from rose and white wines.

### Contributions

These bar plots illustrate the signed _contribution_ of variables from the two data tables. From these figures, we use the direction and the magnitude of these signed contributions to interpret the dimension.

The direction of the signed contribution is the direction of the loadings, and it shows _how_ the variables contribute to the dimension. The variables that contribute in a similar way have the same sign, and those that contribute in an opposite way will have different signs.

The magnitude of the contributions are computed as squared loadings, and they quantify the amount of variance contributed by each variable. Therefore, contribution is similar to the idea of an _effect size_. To identify the important variables, we find the variables that contribute more than average (i.e., with a big enough effect size). When the variables are centered and scaled to have their _sums of squares equals 1_, each variable contributes one unit of variance; therefore, the average contribution is 1/(# of variables of the table).

```{r loadings, fig.height = 4}
res.pls.plot$results.graphs$ctrX.plot
res.pls.plot$results.graphs$ctrY.plot
```

From these two bar plots, the first dimension is characterized by (1) the positive association between `Alcohol` and `Tannin` from the Chemical data and `Woody` and `Astringent` from the Sensory data, and (2) the negative association between these variables and `Hedonic` from the Sensory data.

Together with the latent variable plot, we found that, as compared to the rose and the white wines in the sample, the red wines are less `Hedonic` and stronger in `Alcohol`, `Tannin`, `Woody`, and `Astringent`.

### Circles of correlations

The circle of correlations illustrate how the variables are correlated with each other and with the dimensions. From this figure, the length of an arrow indicates how much this variable is explained by the two given dimensions. The cosine between any two arrows gives their correlation. The cosine between a variable and an axis gives the correlation between that variable and the corresponding dimension. 

In this figure, an angle closer to 0° indicates a correlation close to 1; an angle closer to 180° indicates a correlation close to -1; and an 90° angle indicates 0 correlation. 

```{r corcircles}
res.pls.plot$results.graphs$cirCorX.plot
res.pls.plot$results.graphs$cirCorY.plot
```

These circles of correlations show that `Alcohol`, `Tannin`, `Woody`, `Astringent`, and `Hedonic` are strongly correlated to Dimension 1 with `Henodic` inversely correlated with all other variables. These variables are mostly explained by the first dimension and have close-to-zero correlation with the second dimension (which is not included and discussed in the previous sections).
