---
title: "Wine of The World"
author: "Vincent Guillemot & Hervé Abdi & Soudeh Aderstani Khoubrouy"
date: "`r Sys.Date()`"
output: ioslides_presentation
vignette: >
  %\VignetteIndexEntry{Wine of The World}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style>
.forceBreak { -webkit-column-break-after: always; break-after: column; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, 
  dev = "svg"
)
library(dplyr)
library(knitr)
library(kableExtra)
library(ggplot2)
```

## Load the data

With the following command

```{r winesOf3Colors}
data("winesOf3Colors", package = "data4PCCAR")
```

```{r hidden colors, echo = FALSE}
wineColors <- winesOf3Colors$winesDescriptors$color
wineColors <- as.character(recode(wineColors, red = 'indianred4', 
                white = 'gold', rose = 'lightpink2'))
```


## Argentina {.smaller}

```{r wine argentina, echo = FALSE, results='asis'}
fontsize <- 9.
wines_argentina <- winesOf3Colors$winesDescriptors %>%
    filter(origin == "Argentina")
wines_argentina %>%
    kable("html") %>%
    kable_styling("striped", full_width = F)  %>%
    row_spec(1:nrow(wines_argentina), 
    font_size = fontsize) %>%
    row_spec(0, angle = 0, align = "right", font_size = fontsize)
```

## Canada {.smaller}

```{r wine Canada, echo = FALSE, results='asis'}
wines_canada <- winesOf3Colors$winesDescriptors %>%
    filter(origin == "Canada")
wines_canada %>%
    kable("html") %>%
    kable_styling("striped", full_width = F)  %>%
    row_spec(1:nrow(wines_canada), 
    font_size = fontsize) %>%
    row_spec(0, angle = 0, align = "right", font_size = fontsize)
```

## USA {.smaller}

```{r wine USA, echo = FALSE, results='asis'}
wines_usa <- winesOf3Colors$winesDescriptors %>%
    filter(origin == "USA")
wines_usa %>%
    kable("html") %>%
    kable_styling("striped", full_width = F)  %>%
    row_spec(1:nrow(wines_usa), 
    font_size = fontsize) %>%
    row_spec(0, angle = 0, align = "right", font_size = fontsize)
```

## Four tables

 * Descriptors: origin, color and varietal;
 * Suplementary variables: Price;
 * Chemical data: Acidity, Alcohol, Sugar, and Tannin;
 * Sensory data: fruity, floral, vegetal, spicy, woody, sweet, astringent, and hedonic.

```{r four table, echo = FALSE}
descr <- winesOf3Colors$winesDescriptors %>% 
    select(origin, color, varietal)
suppl <- winesOf3Colors$winesDescriptors %>% 
    select(Price)
chemi <- winesOf3Colors$winesDescriptors %>% 
    select(Acidity, Alcohol, Sugar, Tannin)
senso <- winesOf3Colors$winesDescriptors %>% 
    select(fruity, floral, vegetal, 
           spicy, woody, sweet, astringent, 
           hedonic)
```

## The sensory scores


```{r sensoplot, echo = FALSE, fig.width = 8, fig.height = 5}
senso_long <- tibble(
    yids = rownames(winesOf3Colors$winesDescriptors), 
    as_tibble(scale(senso))) %>%
    tidyr::pivot_longer(
        -yids,
        values_to = "scores",
        names_to = "xids")


(suppressWarnings(heatmap.rxy <- senso_long  %>%
    ggplot(aes(xids, yids, fill = scores)) +
    geom_tile() +
    # geom_text(aes(label = round(scores, 2))) +
    scale_fill_gradient2() +
    theme_bw() +
    theme(
        axis.ticks.x = element_line(color = wineColors),
        axis.text.x = element_text(color = wineColors, size = 7),
    ) +
    labs(x = "", y = ""))) + 
    coord_flip() + 
    theme(legend.position = "bottom")
```


